RadLib: a radiative property model library for CFD

### 2021-09-29 submissions/CPC_response_to_reviewers_revision2.md

3) Line 92 - 95: "Modern band models .. cant lose accuracy when applied to nonhomogeneous media and may still require
   large numbers of RTE evaluations [1]." Should this be "<can> lose accuracy when applied …"?

   **Yes, it should, this typo has been fixed.**

4) In Figure 2, Example S5: Do "n=2" "n=4", etc mean the number of bands in RCSLW? It needs to be clarified.
   Furthermore, can the authors comment on the physical significance of RCSLW n =1 and how is this different from PM
   which uses a spectral average value?

   **The "n=2", etc. refer to the number of gray gases in the RCLSW model. The figure's caption has been updated to clarify this. The figure legend has also been modified to add "n=" to the number listed in Example S2.
The PM model differs from the $n=1$ RCSLW model as follows. The PM model gives a Planck-function-weighted average absorption coefficient, which is not true for the RCSLW model. That statement is now added to the paper in Section 2.1. For n=1, the RCSLW model considers one gray gas as well as a clear gas representing windows in the distribution, discussed on page 4 during the derivation of the gray gas spectral RTE (Eq. 5). 
We have revised the paper to note that the PM model does not include a clear gas. 
The PM model uses a linear mixing of partial-pressure-weighted components for the absorption coefficients, whereas the RCSLW model uses the multiplication method to get the mixture ALBDF from the product of species ALBDF values (Eq. 14), which is then used to compute the absorption coefficient and weight. The method of calculation of the gas absorption coefficient and weight for RCSLW follows from the summary description near the end of section 2.3.**
</font>

<font color='green'>

@@

5) In Figure 2: This is a kind of cosmetic request, but still if the legends and symbols can be made consistent it will
   be easier to read.

   **We have revised the figure as requested. All subplots now use consistent line styles. Figure 3 is also made consistent with Fig. 2.**

@@

6) In lines 499-501: Please clarify the sentence, "The PM absorption coefficient is 27.4 atm−1 m−1 , which results in a
   calculated optical thicknesses of 0.09 and 0.36 m in the thick and thin layers, respectively." Optical thickness is a
   dimensionless quantity, it cannot be 0.36 m. The thicknesses of layers are changing in S2, so what do these 0.09 and
   0.36 refer to?

    **These numbers are the inverse of the PM absorption coefficient times the partial pressures of the radiating gas (CO2) in the two regions and could be considered optical path lengths. We have corrected this issue in the paper and reword to "The PM absorption coefficient is 27.4 (1/atm*m), which results in optical thicknesses greater than 5.5
for the domain sizes considered in Example S2."**
</font>

<font color='green'>

@@

7) In Section 5.1, please clarify which files and functions belong to FDS and which to RadLib. For example, in line
   544, "This was done by editing the radi.f90 file." By looking at the f90 extension the readers can guess that
   radi.f90 is part of FDS and not RadLib. But then in the next sentence, there are references to A_WSGG and KAPPA_WSGG
   functions. Are these part of FDS?

   **This sentence has been revised to clarify which files and functions belong to FDS.**

@@

8) In lines 562-564: "The FDS default model uses a single gray gas with a composition and temperature dependent
   absorption coefficient computed from RADCAL." Is this equivalent to the PM model in RadLib? If so, a comparison with
   PM from RadLib and FDS default will be a good validation. Also, what RTE solver does FDS default use?

   **No, they are not equivalent. FDS uses the RadCal narrow-band model that depends on a user-specified radiative path length with a default value of 0.1 m. This is a geometry-dependent property that is different than a PM coefficient. For the radiative solver, FDS uses a finite-volume discrete ordinates RTE solver. That is noted in the paper in the first paragraph of Section 5.1. We added a note of the path length variable in the revised manuscript.**
</font>

@@

9) In Section 5.2 can the author comment on the computational cost of RadLib-coupled FDS runs? Some quantitative
   information on the computational cost of using RadLib for a coupled run will be very useful. For example, the authors
   can compare the cost of the same FDS configuration simulated without any radiation, with RadLib, and with the
   FDS-default model.

<font color='green'>

10) Line 624: "At present, it includes three major radiation property models - Planck Mean (PM) absorption coefficients
    using the optically thin approximation, the weighted sum of gray gases (WSGG) model,..." PM can be used with any RTE
    solver in principle, not just with optically thin approximation. Is there any limitation within RadLib of using PM
    only with optically thin approximation?

    **This is an error; the reference to the optically thin approximation has been removed. It is true that the PM model can 
be used with any RTE solver. There is no limitation within RadLib that prevents its use with other RTE solvers.**
</font>

### 2021-09-29 manuscript/manuscript.tex
DOL edits mostly done; todo: R3 comment 1, 9

For an absorbing, emitting, and non-scattering medium (considered here), the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$.  Solving the RTE requires information about the radiative properties of the materials and media in the system.  Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model.  Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules and depend on the local thermochemical state of the system (e.g., temperature, pressure, and composition).  Radiation property models typically consist of correlations between radiative properties and high-resolution spectral databases, reducing the huge amount of spectral data and simplifying evaluation of radiation absorption coefficients \cite{Zhang_2002b}.

@@

The Planck Mean (PM) model is the simplest of the three models implemented. This model assumes a single gray gas (with no transparent gas) and computes the absorption coefficient as a function of temperature, pressure, and gas composition.

@@

In general, the PM model performs poorly compared to the WSGG and RCSLW models and the LBL data. The results show that the PM model tends to exaggerate the trends displayed by the other models. The PM data mimics the shape of the other curves in Example S5 as well, but the curve is off-scale and omitted for clarity. For the PM model in Example S5, $Q$ varies from around -80~kW/m$^3$ at $x=0$ to a peak of 200~kW/m$^3$ at x=$1.5$~m and drops to~100 kW/m$^3$ at $x=2$~m.  In Example S2, the PM value of $q(L)/\sigma T^4$ is off the scale at an essentially constant value of unity. The PM absorption coefficient is 27.4 atm$^{-1}$m$^{-1}$, which results in calculated optical thicknesses of 0.09 and 0.36 in the thick and thin layers, respectively. These values are small compared to the domain sizes greater than 0.5 m considered in Example S2.
Example S3 illustrates that the PM model tends to overpredict radiative heat loss at high temperatures, which has been previously observed \cite{Frank_2000,Zhu_2002,Coelho_2002}.

@@

The RadLib model was coupled to the Fire Dynamics Simulator (FDS) code developed at NIST \cite{FDS}. FDS is written in Fortran, and the Fortran interface of Radlib was used for simulations. FDS includes several radiative models, but the spectral WSGG model coupled to a finite volume discrete ordinates model is used here. This was done by editing the \texttt{radi.f90} file in FDS. A \texttt{use rad\textunderscore module} statement is included and the FDS functions \texttt{A\textunderscore WSGG} and \texttt{KAPPA\textunderscore WSGG} are replaced with calls to RadLib's \texttt{get\textunderscore k\textunderscore a\textunderscore oneband}. The \texttt{oneband} version was used because FDS does an outer iteration over the bands/gases, computing properties at a given band for each grid point.

@@

Figure~\ref{f:fds} shows results of the simulation with comparison to experimental data. Three radiation models were included: the FDS default model, and the RCSLW and WSGG models from RadLib (both using four gray gases and one clear gas).
The FDS default model uses a single gray gas with a composition and temperature dependent absorption coefficient computed from RADCAL. Further details are provided in the FDS user guide \cite{FDS}. Figure~\ref{f:fds}a shows radial profiles of mean temperature at two vertical positions. The models produce centerline temperature differences of about 80 K (7\%) between the RCSLW and the WSGG at z=2.5D, and 90 K (12\%) at z=3.5D. The differences are lower at lower points in the flame since the effect of radiation on temperature is cumulative. Figure~\ref{f:fds}b plots height versus radiative emission per unit length along the centerline with the spread in radiative emission around the peak values being 1.3 kW/m (11\%).
%
\begin{figure}
\begin{center}
\begin{tabular}{c c}
(a) & (b) \\
\includegraphics[width=2.5 in]{fig_fds_T_mean.pdf} &
\includegraphics[width=2.5 in]{fig_fds_rad_mean.pdf}
\end{tabular}
\caption{Radial profiles of mean temperature at two heights (a) and height versus radiative emission (b) for FDS simulations of the FM Burner validation case comparing three models to experimental data.}
\label{f:fds}
\end{center}
\end{figure}
%

>>>>

For an absorbing, emitting, and non-scattering medium (considered here), the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$.  Solving the RTE requires information about the radiative properties of the materials and media in the system.  Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model.  Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules and depend on the local thermochemical state of the system (e.g., temperature, pressure, and composition).  
Radiation property models are typically developed as correlations from high-resolution spectral databases, reducing the huge amount of spectral data contained in those databases and simplifying the evaluation of absorption coefficients \cite{Zhang_2002b}.

@@

The Planck Mean (PM) model is the simplest of the three models implemented. This model assumes a single gray gas (with no transparent gas) and computes the absorption coefficient as a function of temperature, pressure, and gas composition. The PM absorption coefficient is a Planck-function-weighted average absorption coefficient.

@@

In general, the PM model performs poorly compared to the WSGG and RCSLW models and the LBL data. The results show that the PM model tends to exaggerate the trends displayed by the other models. The PM data mimics the shape of the other curves in Example S5 as well, but the curve is off-scale and omitted for clarity. For the PM model in Example S5, $Q$ varies from around -80~kW/m$^3$ at $x=0$ to a peak of 200~kW/m$^3$ at x=$1.5$~m and drops to~100 kW/m$^3$ at $x=2$~m.  In Example S2, the PM value of $q(L)/\sigma T^4$ is off the scale at an essentially constant value of unity. 
%
%DOL: old: The PM absorption coefficient is 27.4 atm$^{-1}$m$^{-1}$, which results in calculated optical thicknesses of 0.09 and 0.36 in the thick and thin layers, respectively. These values are small compared to the domain sizes greater than 0.5 m considered in Example S2.
%DOL: revised: The PM absorption coefficient is $k_{PM}=27.4$~atm$^{-1}$m$^{-1}$. The quantity $(k_{PM}P_{CO_2})^{-1}$ is an optical pathlength with values of 0.09~m and 0.36~m in the thick and thin layers, respectively, which are small compared to the domain sizes greater than 0.5~m that are considered in Example S2.
%DOL: preferred, below
The PM absorption coefficient is 27.4~atm$^{-1}$m$^{-1}$, which results in optical thicknesses greater than 5.5
for the domain sizes considered in Example S2.
%

Example S3 illustrates that the PM model tends to overpredict radiative heat loss at high temperatures, which has been previously observed \cite{Frank_2000,Zhu_2002,Coelho_2002}.

@@

The RadLib model was coupled to the Fire Dynamics Simulator (FDS) code developed at NIST \cite{FDS}. FDS is written in Fortran, and the Fortran interface of Radlib was used for simulations. FDS includes several radiative models, but the spectral WSGG model coupled to a finite volume discrete ordinates model is used here. This was done by editing the \texttt{radi.f90} file in FDS. A \texttt{use rad\textunderscore module} statement is included there and the FDS functions \texttt{A\textunderscore WSGG} and \texttt{KAPPA\textunderscore WSGG} are replaced with calls to RadLib's \texttt{get\textunderscore k\textunderscore a\textunderscore oneband}. The \texttt{oneband} version was used because FDS does an outer iteration over the bands/gases, computing properties at a given band for each grid point.

@@

Figure~\ref{f:fds} shows results of the simulation with comparison to experimental data. Three radiation models were included: the FDS default model, and the RCSLW and WSGG models from RadLib (both using four gray gases and one clear gas).
The FDS default model uses a single gray gas with a composition and temperature dependent absorption coefficient computed from RADCAL. The calculation depends on a user-specified radiation path length variable with a default value of 0.1 m, which is used in the FM Burner validation case. Further details are provided in the FDS user guide \cite{FDS}. Figure~\ref{f:fds}a shows radial profiles of mean temperature at two vertical positions. The models produce centerline temperature differences of about 80 K (7\%) between the RCSLW and the WSGG at z=2.5D, and 90 K (12\%) at z=3.5D. The differences are lower at lower points in the flame since the effect of radiation on temperature is cumulative. Figure~\ref{f:fds}b plots height versus radiative emission per unit length along the centerline with the spread in radiative emission around the peak values being 1.3 kW/m (11\%).
%
\begin{figure}
\begin{center}
\begin{tabular}{c c}
(a) & (b) \\
\includegraphics[width=2.5 in]{fig_fds_T_mean.pdf} &
\includegraphics[width=2.5 in]{fig_fds_rad_mean.pdf}
\end{tabular}
\caption{Radial profiles of mean temperature at two heights (a) and height versus radiative emission (b) for FDS simulations of the FM Burner validation case comparing three models to experimental data.}
\label{f:fds}
\end{center}
\end{figure}
%
